HO unification from object language to meta language

### 2024-04-01 paper/main.tex
examples 6

\newcommand{\progressetaleft}{\emph{progress-$\eta$-left}}
\begin{definition}[\progressetaleft]\label{def:progressetaleft}
A link \linketaM{\Gamma}{X}{T} is removed from \linkStore when
$X$ becomes rigid. There are two cases:
\begin{enumerate}
  \item if $X = a$ or $X = f \appsep a_1\ldots a_n$
    we unify the $\eta$-expansion of the $X$ with $T$, that is we run
    $\lambda x.X \appsep x\Ue{} T$  (under
    the context $\Gamma$)
  \item if $X = \lambda x.t$ we run $X \Ue{} T$.
\end{enumerate}
\end{definition}

\newcommand{\progressetaright}{\emph{progress-$\eta$-right}}
\begin{definition}[\progressetaright]\label{def:progressetaright}
A link \linketaM{\Gamma}{X}{T} is removed from \linkStore when
$\testmaybeeta{} T$ does not hold (anymore) and by $\eta$-contracting
$T$ to $T'$ (if possible, else $T' = T$) and executing $X \Ue{} T'$ (under
the context $\Gamma$).
\end{definition}

@@

\newcommand{\progressetadedup}{\emph{progress-$\eta$-deduplicate}}
\begin{definition}[\progressetadedup]\label{def:progressetadedup}
  A link \linketaM{\Gamma}{X_{\vec{s}}}{T} is removed from \linkStore when
  another link \linketaM{\Delta}{X_{\vec{r}}}{T'} is in  \linkStore.
  By \cref{inv:uvaarity} the length of $\vec{s}$ and $\vec{r}$ is the same
  hence we can move the term $T'$ from $\Delta$ to $\Gamma$ by renaming its
  bound variables, i.e. $T'' = T[\vec{r}/\vec{s}]$.
  We then run $X_S \Ue{} T''$ (under the context $\Gamma$).
\end{definition}

We extend the \elpiIn{progress1} predicate
by adding new rules. After passing under all the \elpiIn{abs}
constructors of $l$, there are two cases making a \linketa to progress, 1) \lhs
is instantiated to a rigid term 2) \rhs is no more a \maybeeta or it is a term
which can be reduced to a term with rigid head. 
If \lhs is instantiated to a rigid term $t$, by \cref{prop:nf},
we know that $t$ does not contain any \maybeeta. Let $t'$ the right hand side,
if $t$ is a constant or a function application, then, $t'$, which by
construction has \elpiIn{lam} as head, should be an $\eta$-expansion. We are
therefore allowed to unify $\lambda x. t \appsep x$ (the $\eta$-expanded version
of $t$) with $t'$. Finally, if $t$ is a term with \elpiIn{lam} as head, then
it is not an $\eta$-expansion and therefore, $t$ can be unified with $t'$.

The second way to activate a \linketa is when the \rhs is no more a \maybeeta or
\rhs can be $\eta$-reduced to a term $t$ with rigid head. In both cases, \lhs is
unified with $t'$.

Once a \linketa is activated, it can be removed from \linkStore,
otherwise, the link is kept for a further iteration of \elpiIn{progress}. Note
that this link progression enforce
\cref{prop:nf,inv:linklhs,inv:link-eta-right}: we never commit a term in
the \Ho substitution, since we make unification only when we know that the 
terms are no more \maybeeta, and when \lhs is no more a variable or \rhs is 
no more a \maybeeta, the link is removed from \linkStore.

Another, way to progress \linketa, that we call \linketa deduplication, is when
\linkStore contains two \linketa $l_1$ and $l_2$ with a \lhs having the same
variable address. Let the \lhs of $l_1$ be \elpiIn{uva U R} and the \lhs of
$l_2$ be \elpiIn{uva V S}, then, by \cref{inv:uvaarity}, \elpiIn{R}
and \elpiIn{S} have same scope. Let $t$ be the term obtained by replacing all
each name \elpiIn{S}$_i$ in the \rhs of $l_1$ with \elpiIn{R}$_i$, $t$ is
unified with the \rhs of $l_2$ and one of the two links between $l_1$ and $l_2$
is removed from \linkStore.

A very basic example of \linketa deduplication, is given below:

\printAlll
  {{{\lambda x. X \appsep x, \lambda x. X \appsep x}}}
  {{{E0, E2}}}
  {{{X,A1,1}}}
  {{{,E2,\lambda x.A1_{x}},
    {,E0,\lambda x.A1_{x}}}}

TODO: explain

@@

>>>>

\newcommand{\progressetaleft}{\emph{progress-$\eta$-left}}
\begin{definition}[\progressetaleft]\label{def:progressetaleft}
A link \linketaM{\Gamma}{X}{T} is removed from \linkStore when
$X$ becomes rigid. There are two cases:
\begin{enumerate}
  \item if $X = a$ or $X = y$ or $X = f \appsep a_1\ldots a_n$
    we unify the $\eta$-expansion of the $X$ with $T$, that is we run
    $\lambda x.X \appsep x\Ue{} T$  (under
    the context $\Gamma$)
  \item if $X = \lambda x.t$ we run $X \Ue{} T$.
  \marginpar{do we have the name case in the code?}
\end{enumerate}
\end{definition}

\newcommand{\progressetaright}{\emph{progress-$\eta$-right}}
\begin{definition}[\progressetaright]\label{def:progressetaright}
A link \linketaM{\Gamma}{X}{T} is removed from \linkStore when
$\testmaybeeta{} T$ does not hold (anymore) and by $\eta$-contracting
$T$ to $T'$ (if possible, else $T' = T$) and executing $X \Ue{} T'$ (under
the context $\Gamma$).
\marginpar{there was a note about T' having a rigid head, TO BE FIXED}
\end{definition}

@@

\newcommand{\progressetadedup}{\emph{progress-$\eta$-deduplicate}}
\begin{definition}[\progressetadedup]\label{def:progressetadedup}
  A link \linketaM{\Gamma}{X_{\vec{s}}}{T} is removed from \linkStore when
  another link \linketaM{\Delta}{X_{\vec{r}}}{T'} is in  \linkStore.
  By \cref{inv:uvaarity} the length of $\vec{s}$ and $\vec{r}$ is the same
  hence we can move the term $T'$ from $\Delta$ to $\Gamma$ by renaming its
  bound variables, i.e. $T'' = T'[\vec{r}/\vec{s}]$.
  We then run $T \Ue{} T''$ (under the context $\Gamma$).
\end{definition}

TODO: prove \cref{prop:nf}: we never commit a \maybeeta{} term in
$\sigma$ since we run \Ue{} only when we know that the 
terms are no more \maybeeta, and when \lhs is no more a variable or \rhs is 
no more a \maybeeta, the link is removed from \linkStore.

@@

\paragraph{example 1}

\paragraph{example 2}

\paragraph{example 3}

A very basic example of \linketa deduplication, is given below:
% make test ONLY=7002 TEX=tex
\printAlll
  {{{\lambda x.(X\appsep x),\lambda x.(Y\appsep x)}}}
  {{{A,C}}}
  {{{X,B,1},
    {Y,D,1}}}
  {{{,A,\lambda x.B_{x}},
    {,C,\lambda x.D_{x}}}}

\noindent
The result of $A \Ue{} C$ is that the two \linketa share the same \lhs.
By unifying the two \rhs we get
$\sigma = \{ A~ \mapsto C, B \mapsto D ~\}$.
In turn, given the map $\mathbb{M}$, this second assignment is decompiled to
$\rho = \{~ X \mapsto Y ~\}$ as expected.
